Add Code

### 2025-01-25 code.qmd
Major rewrite
First explain copyright and licenses,
then practice citation and attribution,
then choose own license.

Create file `copyright.qmd` with material
from `choose_license.qmd` and with
additional information on license conditions
and providing attribution.

Create files `data.qmd` and `code.qmd` with
material from `setup.qmd`. Immediately
cover the citation and attribution.

Relicense `Manuscript.qmd` and all code chunks under CC0 1.0. Relicense other
materials under CC BY-SA 4.0.

---
title: "Project Setup"
engine: knitr
---

We will start by setting up a simple example of a reproducible report.

## Create Quarto Project

First, we will need to create a new Quarto project.

If you haven't already, open RStudio -- see @nte-terminal for how to use the terminal instead. Then, click on _File_ > _New Project..._ to open the _New Project Wizard_.

![](images/setup_01.png){width=500px fig-align="center"}

Here, select _New Directory_

![](images/setup_02.png){width=600px fig-align="center"}

And choose the project type _Quarto Project_.

![](images/setup_03.png){width=600px fig-align="center"}

Finally, enter the name of the directory where our report will be created in, for example `code-publishing-exercise`.

As we will use Git to track the version history of files, be sure to check _Create a git repository_. If you don't know what Git is, have a look at the tutorial "[Introduction to version control with git and GitHub within RStudio](https://lmu-osc.github.io/Introduction-RStudio-Git-GitHub/)".

::: {.column-margin}
![renv: A dependency management toolkit for R](images/renv.svg){width=250px}
:::

Also, we will utilize the package [`renv`](https://rstudio.github.io/renv/) to track the R packages our project depends on. Using it makes it easier for others to view and obtain them at a later point in time. Therefore make sure that the box _Use renv with this project_ is checked. Again, if this is the first time you are hearing about `renv`, have a look at the tutorial "[Introduction to {renv}](https://lmu-osc.github.io/introduction-to-renv/)".

If you are already familiar with Markdown and Quarto, you can uncheck the box _Use visual markdown editor_.

![](images/setup_04.png){width=600px fig-align="center"}

Click on _Create Project_. Your RStudio window should now look similar to this:

![](images/setup_05.png){.lightbox fig-align="center" alt="The project `code-publishing-exercise` opened in RStudio. The source pane to the top left has a Quarto file open called \"code-publishing-exercise.qmd\". The console pane to the bottom left indicates by its output that renv is active in the current project. The environment pane to the top right indicates that the environment is currently empty. The output pane to the bottom right shows the files in the current project."}

If, like in the image, a Quarto file with some demo content was opened automatically, you can close and delete it, for example, using RStudio's file manager.

Make sure that your project is in a consistent state according to `renv` by running:

@@

If it reports packages that are not used, synchronize the lock file using:

```{.r filename="Console"}
renv::snapshot()
```

::: {#nte-terminal .callout-note collapse="true"}
### Without RStudio

Without RStudio, one can create a Quarto project with version control and `renv` enabled by typing the following into a terminal:

```{.bash filename="Terminal"}
quarto create project default code-publishing-exercise
cd code-publishing-exercise/
rm code-publishing-exercise.qmd
git init
git checkout -b main
```

Then, one can open an R session by simply typing `R` into the terminal. Next, make sure that `getwd()` indicates that the working directory is `code-publishing-exercise`. Then, initialize `renv`:

```{.r filename="Console"}
renv::init()
```
:::

You are now ready to stage and commit your files. You can either stage files separately or the whole project folder at once. If you do the latter, we recommend you to inspect the untracked changes before staging all of them:

::: {.column-margin}
In file paths, a period (`.`) means "the current directory", while two periods (`..`) mean "the parent directory". Therefore `git add .` means "stage the current directory for committing".
:::

```{.bash filename="Terminal"}
git status
```

Since no commits have been made so far, this should include every file that is not covered by the `.gitignore` file. If everything can be staged for committing -- as is the case in this tutorial -- you can follow up with:

```{.bash filename="Terminal"}
git add .
git commit -m "Initial commit"
```

If you see a file you'd rather not commit, delete it or add its name to the `.gitignore` file. If you don't check your changes before committing, you might accidentally commit something you'd rather not.

::: {#tip-author-identity-unknown .callout-tip}
If `git commit` fails with the message `Author identity unknown`, you need to tell Git who you are. Run the following commands to set your name and email address:

```{.bash filename="Terminal"}
git config user.name "YOUR NAME"
git config user.email "YOUR EMAIL ADDRESS"
```

Then, commit again.
:::

## Decide on Structure

Before adding your project files, it is helpful to decide on a directory structure, that is, how to call each file and where to put it. In general, the directory structure should facilitate understanding a project by breaking it into logical chunks. There is no single best solution, as a good structure depends on where a project's complexity lies. However, it is usually helpful if the different files and folders reflect the execution order. For example, if there are multiple data processing stages, one can possibly differentiate input (raw data), intermediate (processed data), and output files (e.g., figures) and put them into separate folders. Similarly, the corresponding code files (e.g., preparation, modeling, visualization) can be prefixed with increasing numbers.

Luckily, there are already numerous proposals for how to organize one's project files, both general [e.g., @Wilson2017; @TIER4] as well as specific to a particular programming language [e.g., @Marwick2018; @Vuorre2021] or journal [@Vilhuber2021]. We recommend you to follow the standards of your field.

For the purpose of this tutorial, we will provide you with a data set and a corresponding analysis. They are simple enough to be put together in the root folder of your project.

## Add Data

```{r}
#| echo: false

write.csv(palmerpenguins::penguins, file = "data.csv", row.names = FALSE)
```

You can now download the data set we have prepared for you and put it into your project folder: [`data.csv`](data.csv){download=""}

::: {.column-margin}
![palmerpenguins: Palmer Archipelago (Antarctica) Penguin Data](images/palmerpenguins.png){width=250px}
:::

The data set is from the package [`palmerpenguins`](https://allisonhorst.github.io/palmerpenguins/) (v`r packageVersion("palmerpenguins")`) and contains the recorded bill lengths and sex of penguins living on three islands in the Palmer Archipelago, Antarctica. It was made available under the license [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/).

::: {#imp-legal-restrictions .callout-important}
### Consider Legal Restrictions Before Sharing

Everything you put into the project folder will be shared publicly. For reasons of reproducibility, this should include the data you analyze. Of course, you should only share them to the extent you are allowed to, taking into account:

1. applicable privacy laws (e.g., the GDPR for European citizens),
2. contractual obligations (e.g., with your data provider),
3. copyright of the data and their particular structure, and
4. any _sui generis_ database right.

Privacy laws and contractual obligations may require you to create a completely anonymized or synthetic data set^[For example, using [Amnesia](https://amnesia.openaire.eu/), [ARX](https://arx.deidentifier.org/), [sdcMicro](https://sdctools.github.io/sdcMicro/), or [Synthpop](https://www.synthpop.org.uk/).] (if possible), or prohibit any sharing of data, in which case you should provide a reference to a data repository where they can be obtained from. For further information, you can watch the talk "[Data anonymity](https://osf.io/j6fy8)" by Felix Schönbrodt recorded during the LMU Open Science Center Summer School 2023 and have a look at [the accompanying slides](https://osf.io/z6gcu).

Purely factual data such as measurements are usually not copyrightable, but literary or artistic works that cross the threshold of originality are. Additionally, in some jurisdictions data can be subject to _sui generis_ database rights which prevent extracting substantial parts of a database. As a consequence, you need to ensure that you own or have authority to share the data with respect to copyright and similar rights, and to license it to others (see "[Choose a License](choose_license.qmd)").
:::

When distributing a data set, it is important to document the meaning (e.g., units) and valid values of its variables. This is typically done with a _data dictionary_ (also called a _codebook_). In the following, we will demonstrate how to create a simple data dictionary using the R package [`pointblank`](https://rstudio.github.io/pointblank/). You can install it now using:

::: {.column-margin}
![pointblank: Data Validation and Organization of Metadata for Local and Remote Tables](images/pointblank.png){width=250px}
:::

```{.r filename="Console"}
renv::install("pointblank")
```

You can put the code that follows for creating the data dictionary into a new file called `create_data_dictionary.R`.

First, we write down everything we know about the data set. This includes:

- a general description of the data set
- descriptions of all columns
- valid values, where applicable

```{r, filename="create_data_dictionary.R"}
table_info <- c(
  title = "palmerpenguins::penguins",
  description = "Size measurements for adult foraging penguins near Palmer Station, Antarctica"
)
descriptions <- c(
  species = "a character string denoting penguin species",
  island = "a character string denoting island in Palmer Archipelago, Antarctica",
  bill_length_mm = "a number denoting bill length (millimeters)",
  bill_depth_mm = "a number denoting bill depth (millimeters)",
  flipper_length_mm = "an integer denoting flipper length (millimeters)",
  body_mass_g = "an integer denoting body mass (grams)",
  sex = "a character string denoting penguin sex",
  year = "an integer denoting the study year"
)

vals <- list(
  species = c("Adelie", "Gentoo", "Chinstrap"),
  island = c("Torgersen", "Biscoe", "Dream"),
  sex = c("male", "female"),
  year = c(2007, 2008, 2009)
)
```

Depending on the type of data, it may also be necessary to describe measurement instruments, sampling procedures, appropriate weighting, or contact information. In this case, as the data have already been published, we only store a reference to its source:

```{r, class.output = "r code-overflow-wrap", echo = FALSE}

# We have provided the data set as CSV file to the readers.
# Therefore, we cannot assume or require that readers have
# the R package palmerpenguins installed. Instead, we create
# the citation on our end and hide how we obtained it.

dat_source <- citation("palmerpenguins", auto = TRUE) |>
  format(bibtex = FALSE, style = "text")
dat_source |>
  gsub("\\n", " ", x = _) |>
  gsub("\"", "\\\\\"", x = _) |> # Escape any quotation marks
  paste0("dat_source <- \"", ... = _, "\"") |>
  cat()
```

Then, we use `pointblank` to create a data dictionary with this information.

```{r, filename="create_data_dictionary.R"}
vals <- sapply(vals, \(x) {
  paste0(
    "(",
    knitr::combine_words(x, and = " or ", before = "`", after = "`"),
    ")"
  )
})

dat <- read.csv("data.csv")

dict <- pointblank::create_informant(
  dat,
  tbl_name = NA,
  label = table_info[["title"]],
  lang = "en"
) |>
  pointblank::info_tabular(
    Description = table_info[["description"]],
    Source = dat_source
  ) |>
  pointblank::info_columns_from_tbl(stack(descriptions)[2:1]) |>
  pointblank::info_columns_from_tbl(stack(vals)[2:1]) |>
  pointblank::get_informant_report(
    title = "Data Dictionary for `data.csv`"
  )

dict
```

Finally, we can store the data dictionary inside an HTML file and put the HTML file into the project folder as well.

```{.r filename="create_data_dictionary.R"}
pointblank::export_report(dict, filename = "data_dictionary.html")
```

For a more elaborate introduction to `pointblank`, you can read their [Intro to Information Management](https://rstudio.github.io/pointblank/articles/INFO-1.html).

One could go even further by making the information machine-readable in a standardized way. We provide an optional example of that in @nte-frictionless. If you want to learn more about the sharing of research data, have a look at the tutorial "[FAIR research data management](https://lmu-osc.github.io/FAIR-Data-Management/)".

::: {#nte-frictionless .callout-note collapse="true"}
### Create Machine-Readable Variable Documentation

This example demonstrates how the title and description of the data set, the description of the variables and their valid values are stored in a machine-readable way. As before, we also provide a reference to the source.

```{r, class.output = "r code-overflow-wrap", echo = FALSE}

# We have provided the data set as CSV file to the readers.
# Therefore, we cannot assume or require that readers have
# the R package palmerpenguins installed. Instead, we create
# the citation on our end and hide how we obtained it.

citation("palmerpenguins", auto = TRUE)$url |>
  paste0("dat_source <- \"", ... = _, "\"") |>
  cat()
```

Generally, metadata are either stored embedded into the data or externally, for example, in a separate file. We will use the "[frictionless data](https://frictionlessdata.io/)" standard, where metadata are stored separately. Another alternative would be [RO-Crate](https://www.researchobject.org/ro-crate/).

Specifically, one can use the R package [`frictionless`](https://docs.ropensci.org/frictionless/) to create a _schema_ which describes the structure of the data.^[In June 2024, [version 2](https://datapackage.org/) of the frictionless data standard has been released. As of November 2024, the R package `frictionless` only supports the first version, though support for v2 is [planned](https://github.com/frictionlessdata/frictionless-r/labels/datapackage%3Av2).] For the purpose of the following code, it is just a nested list that we edit to include our own information. We also explicitly record in the schema that missing values are stored in the data file as `NA` and that the data are licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/). Finally, the package is used to create a metadata file that contains the schema.

```{.r filename="Console"}
# Read data and create schema
dat_filename <- "data.csv"
dat <- read.csv(dat_filename)
dat_schema <- frictionless::create_schema(dat)

# Add descriptions to the fields
dat_schema$fields <- lapply(dat_schema$fields, \(x) {
  c(x, description = descriptions[[x$name]])
})

# Record valid values
dat_schema$fields <- lapply(dat_schema$fields, \(x) {
  if (x$name %in% names(vals)) {
    modifyList(x, list(constraints = list(enum = vals[[x$name]])))
  } else {
    x
  }
})

# Define missing values
dat_schema$missingValues <- c("", "NA")

# Create package with license info and write it
dat_package <- frictionless::create_package() |>
  frictionless::add_resource(
    resource_name = "penguins",
    data = dat_filename,
    schema = dat_schema,
    title = table_info[["title"]],
    description = table_info[["description"]],
    licenses = list(list(
      name = "CC0-1.0",
      path = "https://creativecommons.org/publicdomain/zero/1.0/",
      title = "CC0 1.0 Universal"
    )),
    sources = list(list(
      title = "CRAN",
      path = dat_source
    ))
  )
frictionless::write_package(dat_package, directory = ".")
```

This creates the metadata file `datapackage.json` in the current directory. Make sure it is located in the same folder as `data.csv`, as together they comprise a [data package](https://specs.frictionlessdata.io/data-package/).
:::

Having added the data and its documentation, one can view and record the utilized packages with `renv`...

```{.r filename="Console"}
renv::status()
renv::snapshot()
```

...and go through the commit routine:

```{.bash filename="Terminal"}
git status
git add .
git commit -m "Add data"
```

## Add Code

In order to have some code which you can practice to share, we have prepared a simple manuscript for you, alongside a bibliography file. The manuscript contains code together with a written narrative. Download the two files to your computer and put them into your project folder.

- [`Manuscript.qmd`](_01_Example/Manuscript.qmd){download=""}
- [`Bibliography.bib`](_01_Example/Bibliography.bib){download=""}

The manuscript explores differences in bill length between male and female penguins, feel free to read through it.

::: {#imp-take-copyright-seriously .callout-important}
### Take Copyright Seriously

If you include work by others in your project -- especially if you intend to make it available publicly --, make sure you have the necessary rights to do so. Only build on existing work for which you receive an express grant of relevant rights. How do you know you are allowed to copy, edit, and share the two files linked above?
:::

:::: {#nte-take-copyright-seriously-solution .callout-note collapse="true"}
### Hint for "Take Copyright Seriously"

Have a look at the [about page](about.qmd).
::::

As the manuscript uses some new packages, install them with:

```{.r filename="Console"}
renv::install()
```

The manuscript also uses the Quarto extension "[apaquarto](https://wjschne.github.io/apaquarto/)", which typesets documents according to the requirements of the American Psychological Association [-@APA2020]. It can be installed in the project using the following command:

::: {#tip-apaquarto .callout-tip}
#### Not a Psychologist?

If you are not a psychologist, you can also skip installing `apaquarto`. If you installed it by accident, run `quarto remove wjschne/apaquarto`.

Note, however, that the file `Manuscript.qmd` we prepared for you uses `apaquarto` by default and you need to set a different `format` in the YAML header if you decide not to use `apaquarto`:

```{.yml filename="Manuscript.qmd"}
format:
  pdf:
    pdf-engine: lualatex
    documentclass: scrartcl
    papersize: a4
```
:::

Also, you need to have a $\TeX$ distribution installed on your computer, which is used in the background to typeset PDF documents. A lightweight choice is [TinyTeX](https://yihui.org/tinytex/), which can be installed with Quarto as follows:

```{.bash filename="Terminal"}
quarto install tinytex
```

You should now be able to render the document using Quarto:

@@

This should create a PDF file called `Manuscript.pdf` in your project folder.

With the code being added, one can use `renv` again to view and record the new packages:

```{.r filename="Console"}
renv::status()
renv::snapshot()
```

::: {#tip-renv-status .callout-tip}
Always run `renv::status()` and resolve any inconsistencies before you commit code to your project. This way, every commit represents a working state of your project.
:::

Finally, make your changes known to Git:

```{.bash filename="Terminal"}
git status
git add .
git commit -m "Add manuscript"
```

@@

- Data may be collected from social network sites using their APIs^[An _application programming interface_ provides the capability to interact with other software using a programming language.] or downloaded from a data repository, or
- an analysis may be conducted with the help of AI providers.

@@

When you start a new session from the project root, the file is automatically read by R and the environment variables can be accessed using `Sys.getenv()`:

@@

### Coding Best Practices

Although we provide the code in this example for you, a few things remain to be said about best practices when it comes to writing code that is readable and maintainable.

- __Use project-relative paths.__ When you refer to a file within your project, write paths relative to your project root. For example, don't write `C:/Users/Public/Documents/my_project/images/result.png`, instead write `images/result.png`.

## Cite Data and Software

If you rely on data or software by others in your research, the question arises whether and how to cite it in your publications.

### Data

Put simply, all data relied upon should be cited to allow for precise identification and access. From the "eight core principles of data citation" by @Starr2015, licensed under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/):

> **Principle 1 – Importance**: "Data should be considered legitimate, citable products of research. Data citations should be accorded the same importance in the scholarly record as citations of other research objects, such as publications."
> 
> **Principle 3 – Evidence**: "In scholarly literature, whenever and wherever a claim relies upon data, the corresponding data should be cited."
> 
> **Principle 5 – Access**: "Data citations should facilitate access to the data themselves and to such associated metadata, documentation, code, and other materials, as are necessary for both humans and machines to make informed use of the referenced data."
> 
> **Principle 7 – Specificity and Verifiability**: "Data citations should facilitate identification of, access to, and verification of the specific data that support a claim. Citations or citation metadata should include information about provenance and fixity sufficient to facilitate verifying that the specific time slice, version and/or granular portion of data retrieved subsequently is the same as was originally cited."

Now, add an appropriate citation for the data set to the manuscript. Does your citation adhere to the principles above?

::: {#nte-cite-palmerpenguins .callout-note collapse="true"}
#### Hint for citing the data set

As the data set is from the R package `palmerpenguins`, one can use the function `citation()` to display a suggested citation:

```{r, echo = -1}
invisible(loadNamespace("palmerpenguins")) # Tell renv that we need this package
citation("palmerpenguins")
```

As this can only be run with the package `palmerpenguins` installed, you can also find a [suggested citation on its website](https://allisonhorst.github.io/palmerpenguins/#citation).

Copy the BibTeX entry to the file `Bibliography.bib` and add an identifier between `@Manual{` and the comma, such that the entry's first line reads `@Manual{horst2020,`. Then, add a sentence to the manuscript such as follows:

```md
The analyzed data are by @horst2020.
```

Render the document to check that the citation is displayed properly.

```{.bash filename="Terminal"}
quarto render Manuscript.qmd
```
:::

### Software

When it comes to software, the answer is a little more nuanced due to the large number of involved dependencies. You can consult @fig-software-citation for general advice whether to cite a particular piece of software or not. As with data, citations should allow for exact identification and access. From the six "software citation principles" by @Smith2016, licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/):

@@

In practice, the first step is to identify all pieces of software the project relies on. A few of them are obvious, such as R itself, Quarto, and the $\TeX$ distribution we installed before. Then there are the individual R packages, Quarto extensions, and $\TeX$ packages. All of them, in turn, may have dependencies and it is up to you decide when not to dig deeper. For example, some R packages are only thin wrappers around other R packages or around system dependencies which also might deserve credit. A system dependency is additional software that you require on your computer apart from the R package.

@@

git status
git add .
git commit -m "Cite data and software"
```

@@

`renv` only records the versions of R packages and of R itself. This means that everything we have not decided to cite in the previous step is not documented anywhere. We will cover system dependencies when [creating a README](make_readme.qmd). For now, however, there is one simple step you can take to record the version of Quarto (and a few other dependencies). Do run the following:

@@

You are now all set up to prepare your project for sharing!

>>>>

---
title: "Add Code"
---

## Add Data Analysis

With the data added to the project folder, you can now analyze them and describe the results. For the purpose of this tutorial, we suggest to conduct a simple analysis that explores whether the bill lengths of penguins vary between sex. You can write your own or use the following example:

::: {#tip-t-test .callout-tip collapse="true"}
### Data Analysis Example

You can copy the following code into the _results_ section of the manuscript:

``````{.qmd .code-overflow-wrap filename="Manuscript.qmd"}
Descriptive statistics of the data set are given in @tbl-descriptive-statistics and individual bill lengths are displayed in @fig-bill-length-comparison.

```{{r}}
#| label: "t-test"
#| echo: false
#| output: "asis"

# Load penguins data set
dat <- read.csv("data.csv")

# Remove rows with NA
dat_clean <- dat[complete.cases(dat), ]

# Treat year as categorical variable
dat_clean$year <- as.factor(dat_clean$year)

# Perform t-test
t_test_result <- t.test(
  bill_length_mm ~ sex,
  data = dat_clean,
  var.equal = FALSE,
  conf.level = 0.95
)

# Describe results
report::report(
  x = t_test_result,
  data = dat_clean,
  rules = "gignac2016"
) |>
  report::as.report_text(summary = TRUE)
```

```{{r}}
#| label: "tbl-descriptive-statistics"
#| echo: false
#| tbl-cap: "Descriptive Statistics"
#| apa-twocolumn: true

report::report_sample(dat_clean, by = "sex", total = FALSE) |>
  knitr::kable()

```

```{{r}}
#| label: "fig-bill-length-comparison"
#| echo: false
#| fig-cap: "Scatter Plot of Bill Lengths by Sex With Violin Plot Showing Quartiles"

# Create ggplot object
ggplot2::ggplot(dat_clean, ggplot2::aes(x = sex, y = bill_length_mm, fill = sex), group = sex) +
  ggplot2::geom_violin(
    draw_quantiles = c(0.25, 0.5, 0.75),
    show.legend = FALSE
  ) +
  ggplot2::geom_jitter(width = 0.15, show.legend = FALSE) +
  ggplot2::labs(x = "Sex", y = "Bill length in mm") +
  ggplot2::theme_minimal()
```
``````

This code uses some new packages, one can use `renv` to view them:

@@

Then, install and record them with:

```{.r filename="Console"}
renv::install()
renv::snapshot()
```
:::

After adding the data analysis, render the manuscript again:

@@

Then, make your changes known to Git:

```{.bash filename="Terminal"}
git status
git add .
git commit -m "Add data analysis"
```

@@

- Data may be collected from social network sites using their APIs^[An **a**pplication **p**rogramming **i**nterface provides the capability to interact with other software using a programming language.] or downloaded from a data repository, or
- an analysis may be conducted with the help of AI providers.

@@

When you start a new session from the project folder, the file is automatically read by R and the environment variables can be accessed using `Sys.getenv()`:

@@

## Style Manuscript

To format manuscripts according to the requirements of a particular journal, [multiple Quarto extensions are available](https://quarto.org/docs/extensions/listing-journals.html). In the following, we will demonstrate the usage of [`apaquarto`](https://wjschne.github.io/apaquarto/), which typesets documents according to the requirements of the American Psychological Association [-@APA2020]. Of course, you may also use an extension for a different journal.

`apaquarto` can be installed in the project folder using the following command:

This downloads the extension into the folder `_extensions` of your project.

Then, change the `format` in the YAML header of your manuscript as follows:

```{.yml filename="Manuscript.qmd"}
format:
  apaquarto-pdf:
    documentmode: man
    keep-tex: true
    pdf-engine: lualatex
    a4paper: true
```

You can now render the manuscript with the new format:

Again, make your changes known to Git:

```{.bash filename="Terminal"}
git status
git add .
git commit -m "Apply custom manuscript style"
```

## Add Code Citation and Attribution

As with data, there are multiple reasons to indicate when using code by others. First, there is software which is actually copied into the own project folder. For example, this is the case for the Quarto extension that you downloaded earlier. These cases matter from a copyright perspective and you need to make sure to have permission for redistribution.

Under which license is `apaquarto` made available?

::: {#tip-apaquarto-license .callout-tip collapse="true"}
### `apaquarto` License (Solution)

According to its GitHub page, `apaquarto` is licensed under CC0\ 1.0. This means that it can be used and distributed even without attribution.
:::

We recommend adding a short paragraph to `LICENSE.txt` to describe its license.

::: {#tip-apaquarto-licensetxt .callout-tip collapse="true"}
### Addition to `LICENSE.txt` (Solution)

```{.txt .code-overflow-wrap filename="LICENSE.txt"}
"apaquarto" stored in "_extensions/wjschne/apaquarto" by W. Joel Schneider available from <https://github.com/wjschne/apaquarto> is licensed under CC0 1.0: <https://creativecommons.org/publicdomain/zero/1.0/>
```

Again, if the license required adding the full license text, you would also need to copy it to `LICENSE.txt` (if not already in there).
:::

Then, there is software that is used (e.g., during the data analysis), but not copied into the project folder. Here, questions of copyright due to redistribution do not apply.

Of course, being transparent about used software for reasons of reproducibility and academic integrity matters in any case. You can consult @fig-software-citation for general advice whether to cite a particular piece of software or not. As with data, citations should allow for exact identification and access. From the six "software citation principles" by @Smith2016, licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/):

@@

In practice, you would identify all pieces of software the project relies on. A few of them are obvious, such as R itself, Quarto, and the $\TeX$ distribution we installed before. Then there are the individual R packages^[By default, `renv` avoids copying them into the project folder. And even that happens (e.g., by using `renv::isolate()`) they are excluded from Git by default.], Quarto extensions, and $\TeX$ packages. All of them, in turn, may have dependencies and it is up to you decide when not to dig deeper. For example, some R packages are only thin wrappers around other R packages or around system dependencies which also might deserve credit. A system dependency is additional software that you require on your computer apart from the R package.

@@

git status
git add .
git commit -m "Add software citation"
```

## Coding Best Practices

Although we provide you with example code in this tutorial, a few things remain to be said about best practices when it comes to writing code that is readable and maintainable.

- __Use project-relative paths.__ When you refer to a file within your project folder, write paths relative to the root (the uppermost folder in your project). For example, don't write `C:/Users/Public/Documents/my_project/images/result.png`, instead write `images/result.png`.

@@

`renv` only records the versions of R packages and of R itself. This means that everything we have not decided to cite or attribute is not documented anywhere. We will cover system dependencies when [creating a `README`](make_readme.qmd). For now, however, there is one simple step you can take to record the version of Quarto (and a few other dependencies). Do run the following:

@@
